---
title: "GIS-R-markdown"
author: "Shane Maguire"
date: "2024-02-29"
output: html_document
self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown
## install packages
```{r packages}
library(rinat)
library(tidyverse)
library(sf)
library(sp)
library(raster)
library(lwgeom)
library(terra)
library(stars)
library(rosm)
library(ggspatial)
library(ggplot2)
library(prettymapr)
library(leaflet)
library(htmltools)
library(leafpop)
library(mapview)
library(dplyr)
```

## read in data
```{R data, echo = TRUE}
hp <- get_inat_obs(taxon_name = "Hyperolius pickersgilli",
                   maxresults = 1000)
```

## filter data for research grade observations only
```{R data-frame, echo = TRUE}
hp <- hp %>% filter(!is.na(latitude) &
                      captive_cultivated == "false" &
                      quality_grade == "research")
```

## transforming data frame into spatial object and defining CRS as Geographic WGS84
```{R spatial, echo = TRUE}
hp <- st_as_sf(hp, coords = c("longitude", "latitude"), crs = 4326)
```

## plotting the spatial data
```{R plot-spatial, echo = TRUE}
ggplot() + geom_sf(data=hp)
```

## adding a base map to spatial plot
```{R plot-basemap, echo = TRUE}
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data=hp)
```

## making interactive map
```{R interactive, echo = TRUE}
leaflet() %>%
  # Add default OpenStreetMap map tiles
  addTiles(group = "Default") %>%  
  # Add our points
  addCircleMarkers(data = hp,
                   group = "Hyperolius pickersgilli",
                   radius = 3, 
                   color = "green") 
```

## adding popup labels 
```{R popup-label, echo = TRUE}
mapview(hp, 
        popup = 
          popupTable(hp,
                     zcol = c("user_login", "captive_cultivated", "url")))
```

## creating live links
```{R live-link}
lhp <- hp %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))

mapview(hp, 
        popup = 
          popupTable(lhp,
                     zcol = c("user_login", "captive_cultivated", "click_url")))
```

## excluding innacurate plots 
```{R exclude}
indices_to_exclude <- c(2, 30, 38)
filtered_hp <- hp %>%
  filter(!row_number() %in% indices_to_exclude)
```

## replotting with selected plots excluded
```{R rerplot, echo = TRUE}
lhp <- filtered_hp %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))

mapview(filtered_hp, 
        popup = 
          popupTable(lhp,
                     zcol = c("user_login", "captive_cultivated", "click_url")))
```
